---
title: "Gapminder Exploration"
data: "September 11, 2018"
output: github_document
---

## Gapminder Dataset

```{r message=FALSE}
library(gapminder)
library(tidyverse)
```

Below is gapminder dataset based on UN statistics in different countries. 

```{r}
gapminder
head(gapminder)
tail(gapminder)
summary(gapminder)
```

## LifeExp, Pop and gdpPercap Information

Below is lifeExp, pop and gdpPercap information between year 1952 and year 2007 

```{r}
summary(gapminder$lifeExp)
summary(gapminder$pop)
summary(gapminder$gdpPercap)
```

## Making some plots 

Here is a plot of lifeExp vs year based on UN statistics from year 1952 to 2007

```{r echo=FALSE}
plot(lifeExp~year, gapminder)
title(main="lifeExp vs year")
```

Here is a plot of lifeExp vs gdpPercap based on UN statistics from year 1952 to 2007, we can see the correlation between lifeExp and economy

```{r echo=FALSE}
plot(lifeExp~gdpPercap, gapminder)
title(main="lifeExp vs gdpPercap")
```

Below is a plot of lifeExp vs gdpPercap by using a log scale plot of gdpPercap. 

```{r echo=FALSE}
plot(lifeExp~log(gdpPercap), gapminder)
title(main="lifeExp vs gdpPercap with log scale of gdpPercap")
```

After log scaling, we can find that there is a linear relationship between lifeExp and gapPercap, showing that people in rich countries have long lifeExp. 

Next, we will plot lifeExp vs gdpPercap on diffrent continents

```{r echo=FALSE}
ggplot(gapminder,aes(x=log(gdpPercap),y=lifeExp))+geom_point()+geom_smooth(se=FALSE)+facet_wrap(~continent)
```

Finally, a plot of lifeExp vs year on different continents is shown below

```{r echo=FALSE}
ggplot(gapminder,aes(x=year,y=lifeExp))+geom_point()+facet_wrap(~continent)
```
